---
title: "data_cleaning Grand Falls"
author: "Luca Comba"
date: "2/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(skimr)
```

Daily Observation Table for January 1 2000 to January 14 2021
Flag Definition Line: 

- M - Missing; 
- E - Estimated; 
- N/A - Not Available

Variables:

- Station Name
- Latitude (deg)
- Longitude (deg)
- Elevation (ft)
- Year
- Month
- Day
- Avg Temp
- Avg Temp Flag
- Penman PET (mm) [stands for potential evapotranspiration and is defined as the amount of evaporation + transpiration that would occur if a sufficient water source were available.]
- Rainfall (mm)
- Rainfall flag

```{r}
meteo_grand <- read.csv("C:/Users/Luca/Desktop/SPRING2021/Statistical Research/CSV Cleaned/Meteorological Data for Grand Forks_cleaned.csv", encoding="UTF-8")

summary(meteo_grand)
skim(meteo_grand)
```


We are replacing empty with the Not Available value

```{r}
meteo_grand[meteo_grand==""]<-NA

# check how many NA each variable has
for (i in 1:13) {
  print(c(colnames(meteo_grand[i]), sum(is.na(meteo_grand[,i]))))
}
```

then we want to see what type of variable we have

```{r}
glimpse(meteo_grand)
```

We can see that the integers variables are *Elevation*, *Year*, *Month* and *Day*.


**Univariate Outliers**
We will use boxplots and z-scores for finding outliers.

- Average Temperature outliers
```{r}
# Average Temperature
zScores_avg_temp_grand <- (meteo_grand$Avg.Temp - mean(meteo_grand$Avg.Temp, na.rm = T)) / sd(meteo_grand$Avg.Temp, na.rm = T)
sum(abs(zScores_avg_temp_grand)> 3.29, na.rm = TRUE)
```
We do not have any outliers for the `average temeperature`
In fact in the boxplot there are no outliers. The histogram shows us that the distribution of `average temperatures` is almost normally distributed.
```{r}
hist(meteo_grand$Avg.Temp)
boxplot(meteo_grand$Avg.Temp)
```

some temperatures have been estimated

- rainfall outliers

```{r}
# Rainfall
zScores_rainfall_grand <- (meteo_grand$Rainfall - mean(meteo_grand$Rainfall, na.rm = T)) / sd(meteo_grand$Rainfall, na.rm = T)
sum(abs(zScores_rainfall_grand)> 3.29, na.rm = TRUE)
```
With the zScore we have found that there are 96 outliers. We need to remember that out of 7685 cases we have 2804 `NA` because those data were missing and 125 were estimated.

```{r}
sum(is.na(meteo_grand$Rainfall))

sum(meteo_grand["Rainfall.Flag"]=="E", na.rm = TRUE)
sum(meteo_grand["Rainfall.Flag"]=="M", na.rm = TRUE)
```
The histogram is highly skewed right.

```{r}
hist(meteo_grand$Rainfall)
boxplot(meteo_grand$Rainfall)
```

- Penman PET (in mm)
```{r}
# Penman
zScores_penman_grand <- (meteo_grand$Penman.PET - mean(meteo_grand$Penman.PET, na.rm = T)) / sd(meteo_grand$Penman.PET, na.rm = T)
sum(abs(zScores_penman_grand)> 3.29, na.rm = TRUE)
```
For the variable Penman PET we have 7 outliers.

Also the histogram of Penman is skewed right.

```{r}
hist(meteo_grand$Penman.PET)
boxplot(meteo_grand$Penman.PET)
```

**Bivariate Outliers**
We are going to use scatter plots and mahalanobis distance.

- Average Temperature with Rainfall
```{r}
mahaDistance = mahalanobis(cbind(meteo_grand$Avg.Temp, meteo_grand$Rainfall), center = c(mean(meteo_grand$Avg.Temp, na.rm = T), mean(meteo_grand$Rainfall, na.rm = T)),cov = cov(cbind(meteo_grand$Avg.Temp, meteo_grand$Rainfall), use = "complete.obs"))

#mahaDistance
sum(pchisq(q = mahaDistance, df = 1, lower.tail = FALSE) >= 0.05, na.rm = T)

leverage = mahaDistance / (nrow(meteo_grand) -1) + 1 / nrow(meteo_grand)

sum(leverage > mean(leverage)*3, na.rm = T)
```
- Average Temperature with Penman.PET
```{r}
mahaDistance = mahalanobis(cbind(meteo_grand$Avg.Temp, meteo_grand$Penman.PET), center = c(mean(meteo_grand$Avg.Temp, na.rm = T), mean(meteo_grand$Penman.PET, na.rm = T)),cov = cov(cbind(meteo_grand$Avg.Temp, meteo_grand$Penman.PET), use = "complete.obs"))

#mahaDistance
sum(pchisq(q = mahaDistance, df = 1, lower.tail = FALSE) >= 0.05, na.rm = T)

leverage = mahaDistance / (nrow(meteo_grand) -1) + 1 / nrow(meteo_grand)

sum(leverage > mean(leverage)*3, na.rm = T)
```


- Rainfall with Penman.PET
```{r}
mahaDistance = mahalanobis(cbind(meteo_grand$Rainfall, meteo_grand$Penman.PET), center = c(mean(meteo_grand$Rainfall, na.rm = T), mean(meteo_grand$Penman.PET, na.rm = T)),cov = cov(cbind(meteo_grand$Rainfall, meteo_grand$Penman.PET), use = "complete.obs"))

#mahaDistance
sum(pchisq(q = mahaDistance, df = 1, lower.tail = FALSE) >= 0.05, na.rm = T)

leverage = mahaDistance / (nrow(meteo_grand) -1) + 1 / nrow(meteo_grand)

sum(leverage > mean(leverage)*3, na.rm = T)
```
- Rainfall with Penman.PET and Temperature
```{r}
mahaDistance = mahalanobis(cbind(meteo_grand$Rainfall, meteo_grand$Penman.PET, meteo_grand$Avg.Temp), center = c(mean(meteo_grand$Rainfall, na.rm = T), mean(meteo_grand$Penman.PET, na.rm = T), mean(meteo_grand$Avg.Temp, na.rm = T)),cov = cov(cbind(meteo_grand$Rainfall, meteo_grand$Penman.PET, meteo_grand$Avg.Temp), use = "complete.obs"))

#mahaDistance
sum(pchisq(q = mahaDistance, df = 2, lower.tail = FALSE) >= 0.05, na.rm = T)

leverage = mahaDistance / (nrow(meteo_grand) -1) + 1 / nrow(meteo_grand)

sum(leverage > mean(leverage)*3, na.rm = T)
```

- Every variable for finding bivariate outliers
```{r}
n = ncol(meteo_grand)

#for(j in 1:(n-1)){
#    for(i in 1:(n-j)){
    
      # print name of variable
      #print(c(colnames(meteo_grand[i]), colnames(meteo_grand[j])))
      
      
      # Calculate mahaDistance
      #mahaDistance = mahalanobis(cbind(meteo_grand[,i], meteo_grand[,j]), center = c(mean(meteo_grand[,i], na.rm = T), mean(meteo_grand[,j], na.rm = T)),cov = cov(cbind(meteo_grand[,i], meteo_grand[,j]), use = "pairwise.complete.obs"))
    
      #sum(pchisq(q = mahaDistance, df = 1, lower.tail = FALSE) >= 0.05, na.rm = T)
      
      
      
      # Calculate leverage
      #leverage = mahaDistance / (nrow(meteo_grand) -1) + 1 / nrow(meteo_grand)
      
      #sum(leverage > mean(leverage)*3, na.rm = T)
  #}
#}

#for (i in 1:12) {
  #for (j in (i+1):13) {
    #print(c(i,j))
    
    #mahaDistance = mahalanobis(cbind(meteo_grand[,i], meteo_grand[,j]), center = c(mean(meteo_grand[,i], na.rm = T), mean(meteo_grand[,j], na.rm = T)),cov = cov(cbind(meteo_grand[,i], meteo_grand[,j]), use = "pairwise.complete.obs"))
    
  #}
#}



```


**Skewness and kurtosis**
```{r}
require(moments)

# for average temperature
skewness(meteo_grand$Avg.Temp)
kurtosis(meteo_grand$Avg.Temp)
pnorm(q =  (skewness(meteo_grand$Avg.Temp) - 0) / sqrt(6 / length(meteo_grand$Avg.Temp)), lower.tail = FALSE)
pnorm(q =  (kurtosis(meteo_grand$Avg.Temp) - 0) / sqrt(24 / length(meteo_grand$Avg.Temp)), lower.tail = FALSE)
```

```{r}
# for rainfall
skewness(meteo_grand$Rainfall, na.rm = T)
kurtosis(meteo_grand$Rainfall, na.rm = T)
pnorm(q =  (skewness(meteo_grand$Rainfall, na.rm = T) - 0) / sqrt(6 / length(meteo_grand$Rainfall)), lower.tail = F)
pnorm(q =  (kurtosis(meteo_grand$Rainfall, na.rm = T) - 0) / sqrt(24 / length(meteo_grand$Rainfall)), lower.tail = F)

```

```{r}
# for penman.PET
skewness(meteo_grand$Penman.PET)
kurtosis(meteo_grand$Penman.PET)

pnorm(q =  (skewness(meteo_grand$Penman.PET) - 0) / sqrt(6 / length(meteo_grand$Penman.PET)), lower.tail = FALSE)
pnorm(q =  (kurtosis(meteo_grand$Penman.PET) - 0) / sqrt(24 / length(meteo_grand$Penman.PET)), lower.tail = FALSE)
```

**Bivariate homoscedasticity**
scatter plots are usually enough

```{r}
pairs(meteo_grand[,c(8,10,12)])
```

**Bivariate linearity**
use a qq-plot, it's a bit cleaner than just an outright scatterplot - and be ready to transform anything that isn't linear
```{r}
qqplot(meteo_grand$Avg.Temp, meteo_grand$Rainfall)
qqplot(meteo_grand$Avg.Temp, meteo_grand$Penman.PET)
qqplot(meteo_grand$Rainfall, meteo_grand$Penman.PET)
```

**Multicollinearity & Singularity**
```{r}
corMat_grand = cor(cbind(meteo_grand$Avg.Temp, meteo_grand$Rainfall, meteo_grand$Penman.PET), use = "complete.obs")
corMat_grand
eig = eigen(corMat_grand)
```
There appears to be no multicollinearity or singularity because none of the correlations are above 0.9.

```{r}
eig = eigen(corMat_grand)
sqrt(max(eig$values)/eig$values)
```
There appears to be no multicollinearity or singularity because none of the condition indices isabove 30


