---
title: "Research visualizations for paper"
author: "Luca Comba"
date: "4/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

Using R it will be convinent to create graph and visualization for the paper. It is our intent to utilize R packages such as `ggplot2` to visualize data and models to export and include them into the written and presentation of the research paper.

# For the paper

The paper is structured with 5 sections. They all base their content on the dataset and analysis made on the dataset. It is important to follow readability structure of the data and be mindful of colors and labels in order to create easy to understand and read visualizations.

We will start by loading the needed r packages.

```{r, results='hide', warning=FALSE, results='hide',message=FALSE}
library(tidyverse) # this includes ggplot and dplyr
library(maps) # to create maps
library(rgdal) # for reading shape files
library(sf)
library(grid) # for making grids
library(gridExtra) # another package for tables
library(cowplot) # for making table of plots
library(RColorBrewer) # for nice color pallet
```

It will be necessary to load the correct dataset.

```{r, results='hide'}
# data for Fargo
fargo <- read.csv("C:/Users/Luca/Desktop/SPRING2021/Statistical Research/Correct Dataset/fargo_corrected_discharge.csv", encoding="UTF-8")
# data for Grand
grand <- read.csv("C:/Users/Luca/Desktop/SPRING2021/Statistical Research/Correct Dataset/grand_corrected_discharge.csv", encoding="UTF-8")
```

```{r}
# reading different river shape coordinates data
shape_river <- readOGR('C:/Users/Luca/Desktop/SPRING2021/Statistical Research/RRB_Files', 'RedRiver')

shape_basin <- readOGR('C:/Users/Luca/Desktop/SPRING2021/Statistical Research/RRB_Files', 'RedRiverBasin')
```

## 1. Introduction

For the introduction we would like to show a map of the Red River region and show the two weather station that have been collecting the data.

```{r}
points_cities <- data.frame (lat = c(46.897, 47.83633),
                  lon = c(-96.812, -97.066838))
# ggplot
map_general <- map_data("state") %>%
  ggplot() + 
  geom_polygon(aes(long, lat, group = group), fill = 'gray90', color = 'black') +
  geom_point(data = points_cities, aes(lon,lat), color = "red") +
  annotate("text", x = -97.1, y = 47.1, label = "Fargo") +
  annotate("text", x = -97, y = 48, label = "Grand Forks") +
  coord_map() + 
  coord_cartesian(xlim= c(-100, -93), ylim = c(45, 49)) +
  theme_void()
map_general
# saving
```

With Red River in the map

```{r}
# ggplot
map_red_river <- map_data("state") %>%
  ggplot(aes(long, lat)) + 
  geom_polygon(aes(group = group),fill = 'gray90', color = 'black') +
  geom_path(aes(group = group), data = shape_river, color = 'steelblue2') +
  geom_point(data = points_cities, aes(lon,lat), color = "red") +
  annotate("text", x = -97.1, y = 47.1, label = "Fargo") +
  annotate("text", x = -97, y = 48, label = "Grand Forks") +
  coord_map() + 
  coord_cartesian(xlim= c(-100, -93), ylim = c(45, 49)) +
  theme_void()
map_red_river
```

A more complete map

```{r}
# ggplot
map_red_river <- map_data("state") %>%
  ggplot(aes(long, lat)) + 
  geom_polygon(aes(group = group), fill = 'gray90', color = 'black') +
  geom_polygon(aes(group = group), data = shape_basin, color = 'brown', fill = "burlywood1", alpha = 0.5) +
  geom_path(aes(group = group), data = shape_river, color = 'steelblue2') +
  geom_point(data = points_cities, aes(lon,lat), color = "red") +
  annotate("text", x = -97.1, y = 47.1, label = "Fargo") +
  annotate("text", x = -97, y = 48, label = "Grand Forks") +
  coord_map() + 
  coord_cartesian(xlim= c(-100, -93), ylim = c(45, 49)) +
  ggtitle("Map of Red River Basin", subtitle = "") +
  xlab("longitude") + ylab("latitude") +
  theme_minimal()
map_red_river
```


Source: [https://stackoverflow.com/questions/51333546/how-to-plot-rivers-efficiently/51334108](https://stackoverflow.com/questions/51333546/how-to-plot-rivers-efficiently/51334108)

```{r}
shapeUFs <- readOGR('C:/Users/Luca/Desktop/SPRING2021/Statistical Research/RRB_Files', 'RedRiverBasin')
shapeHid <- readOGR('C:/Users/Luca/Desktop/SPRING2021/Statistical Research/RRB_Files', 'RedRiver') 

ggplot(shapeUFs, aes(long, lat, group = group)) +
  geom_polygon(fill = 'gray90', color = 'black') +
  geom_path(data = shapeHid, color = 'steelblue2') +
  coord_map() + theme_void()
```